\documentclass[a4paper,12pt]{report}
% document customization
\usepackage[margin=1in]{geometry}
% load custom font
\usepackage{fontspec}
\setmainfont{Carlito}
% code/monospace
\usepackage{listings}
% ability to use \FloatBarrier
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{tabularx}
\graphicspath{{./assets/}}
\usepackage{wrapfig}
% charts
\usepackage{bchart}
% href links
\usepackage[hidelinks]{hyperref}
\hypersetup{
    colorlinks,
    allcolors=black
}
% glossaries
% add 'nomain' if no main glossary exists
\usepackage[nomain,acronym,nopostdot]{glossaries}
\input{glossaries.tex}
\makeglossaries
% remove glossary section titles
\renewcommand{\glossarysection}[2][]{}
\usepackage{bookmark}
% create bibliography
% run 'biber <tex name>' to convert .bib files
\usepackage[style=authoryear-ibid,backend=biber,natbib=true]{biblatex}
\addbibresource{refs.bib}
\addbibresource{web-refs.bib}
% section customization
\usepackage{titlesec}
\titleformat{\chapter}{\normalfont\huge}{\thechapter.}{20pt}{\huge\it}
\titlespacing*{\chapter}{0pt}{-60pt}{20pt}
% caption customization
\usepackage[labelfont=bf, skip=5pt, font=small]{caption}
% paragraph customization
\renewcommand{\baselinestretch}{1.25}
\setlength{\parskip}{1em}
\setlength{\parindent}{0em}


\begin{document}

\begin{titlepage}
    \rightline{\textbf{FAO} Allan Callaghan}

	\begin{center}
		\textbf{...}

		\textbf{BSc (Hons.) Computer Science}

		\textbf{2022-2023}

		\textbf{IP 40}

		\vspace{4ex}

		\textbf{Title:} <Your IP 40 Title>

        \textbf{Author:} Leo Spratt

        \textbf{Supervisor:} Allan Callaghan

        \textbf{E-mail:} ls959@canterbury.ac.uk

		\vspace{4ex}

        \small{This report is submitted in partial fulfilment of the requirement for the BSc in \textsl{Computer Science} at Canterbury Christ Church University

		I declare that this report is my own original work containing no personal data as defined in the Data Protection Act (1998) and that I have read, understood and accept the University's regulations on plagiarism/intellectual property rights/research ethics (in particular the Research Governance Handbook) and the IP 40 Module Handbook.

        Finally, I accept that no copy of my Individual Project 40 will ever be returned regardless of the circumstances.}

		Signed: \verb|              |

		Date Of Submission: 5th May 2023
	\end{center}
\end{titlepage}

% \nocite{*} % add any unsighted references to bibliography

\begin{abstract}
	\input{chapters/abstract.tex}
\end{abstract}
\chapter*{Acknowledgements}
\input{chapters/acknowledgements.tex}
\tableofcontents
\newpage
\listoffigures
\newpage
\listoftables
\newpage
\lstlistoflistings
\newpage
\chapter{Introduction}
\input{chapters/intro.tex}
\chapter{Legal Considerations}
There are no legal considerations.
\chapter{Ethical Considerations}
There are no ethical considerations.
% START main chapters
\chapter{Existing Solutions}
\input{chapters/main/existing_solutions.tex}
\chapter{Research}
\input{chapters/main/research.tex}
\chapter{Prototyping}
\input{chapters/main/prototypes.tex}
\chapter{Domain Analysis}
\input{chapters/main/domain.tex}
% END main chapters
\chapter{Conclusion}
\input{chapters/conclusion.tex}

% use \cite{} or \parencite{} cite a reference
\printbibliography[heading=bibintoc]

\appendix
\chapter{Glossary}
\printglossaries
\chapter{Marking Scheme}
\chapter{Changes to the Project Initiation Document}
\chapter{Current Environment Investigation Report}
Not Applicable.
\chapter{Requirements Specification}
Not Applicable.
\chapter{Design Report}
Not Applicable.
\chapter{Implementation}
Not Applicable.
\chapter{Testing}
\input{chapters/appendices/testing.tex}
\chapter{User Guide}
\input{chapters/appendices/user-guide.tex}
\chapter{Project Management}
\input{chapters/appendices/project-management.tex}
\chapter{Meetings With Supervisor}
\input{chapters/appendices/meetings.tex}
\chapter{Agile Development: Timebox 1}
Not Applicable.
\chapter{Agile Development: Timebox 2}
Not Applicable.
\chapter{Agile Development: Timebox 3}
Not Applicable.
\chapter{Diagrams}
\input{chapters/appendices/diagrams.tex}
\chapter{Prototype Designs}
\input{chapters/appendices/designs.tex}
\chapter{Links}
\input{chapters/appendices/links.tex}

\glsaddallunused
\end{document}
